---
output:
  word_document: default
  html_document: default
---
# MIS503 Finla Project
## Weppler, Alex
### Zillow Home Value Index Analysis

### Wake County Home Sales 
a. What have been the overall trends in Wake County Home Values?
    
 It seems that the overall trends are almost cyclical, with moments of up and down. Overall, values have steadily risen, but there have been low periods throughout. 
    
b. There were dips in home values in the past 20 years. What years did these occur?

 There was a dip 1997- 1999 and also 2008-2012.
 
c. Based on the analysis, where would be the least expensive area to purchase home? Most expensive
area?

 The least expensive area to buy a home would be Zebulon. The most expensive area to buy a house would be Apex.
 
d. Are any area home values trending down? Is there one area that stands out compared to others?

  Yes, while everyone had an overall increase, you can tell that some areas, while competitive to other areas on the scatter plot in the 90"s, have dropped value in comparison into the late 2000's. The most noticable one seems to be Raleigh (the blue is very hard to differentiate, I think its Raleigh). Its home values were comparable to Knightsdale, Mooresville, and others. In recent years, those values have grown exponentially, while Raleigh is only remaining a constant with inflation trends. 

```{r}
library(tidyverse)
library(readr)
SingleFamilyResidenceSales <- read_csv("SingleFamilyResidenceSales.csv")
```

```{r}
WakeCountySales<- SingleFamilyResidenceSales %>%
  filter(State %in% "NC") %>%
  filter(CountyName %in% "Wake County") %>%
  select("RegionName", "State", "CountyName", "Metro", "1996-05", "1997-05", "1998-05", "1999-05", "2000-05", "2001-05", "2002-05", "2002-05", "2003-05", "2004-05","2005-05", "2006-05", "2007-05", "2008-05", "2009-05", "2010-05", "2011-05", "2012-05", "2013-05", "2014-05", "2015-05", "2016-05", "2017-05", "2018-05") %>%
  rename("1996" = "1996-05") %>%
 rename("1997" = "1997-05", "1998"="1998-05") %>% 
  rename("1999" = "1999-05")

WakeCountySales<- WakeCountySales%>%
  rename("2000" = "2000-05") %>%
  rename("2001" = "2001-05") %>%
  rename("2002" = "2002-05", "2003" = "2003-05") %>%
  rename("2004" = "2004-05", "2005" = "2005-05") %>%
  rename("2006" = "2006-05", "2007" = "2007-05", "2008" = "2008-05", "2009" = "2009-05", "2010" = "2010-05", "2011" = "2011-05", "2012" = "2012-05", "2013" = "2013-05", "2014" = "2014-05", "2015" = "2015-05", "2016" = "2016-05", "2017" = "2017-05", "2018" = "2018-05")
  
WakeCountySales<- pivot_longer(WakeCountySales, cols= c("1996","1997", "1998", "1999","2000","2001", "2002", "2003", "2004","2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018"), names_to = "YR", values_to = "ZHVI")

ggplot(WakeCountySales, aes(YR, ZHVI, color = RegionName)) +
  geom_point() +
  labs( title = "Wake County Home Sales", x= "Year", y= "Home Value (In $)") + 
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)

ggplot(WakeCountySales, aes(YR, ZHVI, color = RegionName)) +
  geom_col() +
  labs( title = "Wake County Home Sales", x= "Year", y= "Home Value (In $)") + 
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)
```


### NC Rental Market

a. What has been the overall trend in the rental market around the state? Are there any cities that
have not followed this trend?

The general trend is upward. It seems that pretty much all cities have increasing rent, except for Fayetteville, which has rermained steady. 

b. Where is the most expensive city to rent in? Least expensive?

Most expensive is Asheville, least expensive is Fayetteville.

c. You are trying decide between Wilmington and Asheville. Which market has the lowest rent?

Wilmington has lower rent and less hipsters, it is the obvious choice. 

```{r}
SingleFamilyResidenceRental <- read_csv("SingleFamilyResidenceRental.csv")
```

```{r}
Rentals<- SingleFamilyResidenceRental %>%
  filter(State %in% "NC") %>%
  filter(RegionName %in% c("Asheville", "Charlotte", "Durham", "Fayetteville", "Raleigh", "Wilmington")) %>%
  select("RegionName","State", "2010-11", "2011-11", "2012-11", "2013-11", "2014-11", "2015-11", "2016-11", "2017-11", "2018-10")

Rentals<- Rentals %>%
  rename("2010"= "2010-11","2011"= "2011-11","2012"= "2012-11","2013"= "2013-11","2014"= "2014-11","2015"= "2015-11","2016"= "2016-11","2017"= "2017-11","2018"= "2018-10")

Rentals<- pivot_longer(Rentals, cols= c("2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018"), names_to = "YR", values_to = "ZHVI") 

ggplot(Rentals, aes(YR, ZHVI, color = RegionName)) +
  geom_point() +
  labs( title = "Rental Costs in NC Cities", x= "Year", y= "Average Rental Cost") + 
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Average rental Cost", labels = scales::comma)

ggplot(Rentals, aes(YR, ZHVI, color = RegionName)) +
  geom_col() +
  labs( title = "Rental Costs in NC Cities", x= "Year", y= "Average Rental Cost") + 
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Average rental Cost", labels = scales::comma)
```

### Home Values in Select Rental Markets

a. According to the results, which market has the lowest median price (represented as horizontal bar in
box plot)?
 
 Charlotte-Concord-Gastonia

b. The violin plot will show density meaning the wider the plot is, the more observations occur within
that area. Which market has the most density around the median value of homes?

  Charlotte-Concord-Gastonia

```{r}
NCHomeSales<- SingleFamilyResidenceSales %>%
  filter(State %in% "NC") %>%
  select("State", "Metro", "1996-05", "1997-05", "1998-05", "1999-05", "2000-05", "2001-05", "2002-05", "2002-05", "2003-05", "2004-05","2005-05", "2006-05", "2007-05", "2008-05", "2009-05", "2010-05", "2011-05", "2012-05", "2013-05", "2014-05", "2015-05", "2016-05", "2017-05", "2018-05")
NCHomeSales<- filter(NCHomeSales, Metro %in% c("Asheville", "Charlotte-Concord-Gastonia", "Raleigh", "Wilmington"))
NCHomeSales<- NCHomeSales %>%
  rename("1996" = "1996-05","1997" =  "1997-05","1998" =  "1998-05", "1999" = "1999-05","2000" = "2000-05","2001"= "2001-05","2002" = "2002-05","2003"= "2003-05","2004"= "2004-05","2005" = "2005-05","2006" = "2006-05", "2007" = "2007-05","2008" = "2008-05","2009" = "2009-05","2010" = "2010-05","2011" = "2011-05","2012" = "2012-05","2013" = "2013-05","2014" = "2014-05", "2015"= "2015-05", "2016"= "2016-05", "2017" = "2017-05", "2018" = "2018-05")

NCHomeSales<- pivot_longer(NCHomeSales, cols= c("1996","1997", "1998", "1999","2000","2001", "2002", "2003", "2004","2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018"), names_to = "YR", values_to = "ZHVI")

NCHomeSales<- group_by(NCHomeSales, Metro)

ggplot(NCHomeSales, aes(Metro, ZHVI)) +
  geom_violin() +
  geom_boxplot(width=.1)
```

### Relocation Home Value Comparison

a. Based on your analysis, which city’s housing is most affordable? Least affordable?

Most affordable is Houston- Woodlands- Sugar land area. Least affordable is New York- Newark- New Jersey.

b. Which cities saw the largest change in prices over the past 5 years? Which city has remained more
consistent (i.e., no huge swings up or down in home values)?

New York- Newark- New Jersey has had the largest change. Houston- Woodlands- Sugar land has been the most consistent.  

c. During the market downturn in 2012, which cities were most impacted? Which cities have
recovered?

Chicago and New York and had the biggest impact. Chicago is slowly recovering, whereas New York has recovered and exceeded previous values. 

```{r}
NationalHomeSales<- SingleFamilyResidenceSales %>%
  filter(State %in% c("IL","NY","CO","TX")) %>%
  filter(RegionName %in% c("Chicago","Denver","Houston", "New York")) %>%
  select("State", "Metro","RegionName", "1996-05", "1997-05", "1998-05", "1999-05", "2000-05", "2001-05", "2002-05", "2002-05", "2003-05", "2004-05","2005-05", "2006-05", "2007-05", "2008-05", "2009-05", "2010-05", "2011-05", "2012-05", "2013-05", "2014-05", "2015-05", "2016-05", "2017-05", "2018-05") 

NationalHomeSales <- NationalHomeSales %>%
   rename("1996" = "1996-05","1997" =  "1997-05","1998" =  "1998-05", "1999" = "1999-05","2000" = "2000-05","2001"= "2001-05","2002" = "2002-05","2003"= "2003-05","2004"= "2004-05","2005" = "2005-05","2006" = "2006-05", "2007" = "2007-05","2008" = "2008-05","2009" = "2009-05","2010" = "2010-05","2011" = "2011-05","2012" = "2012-05","2013" = "2013-05","2014" = "2014-05", "2015"= "2015-05", "2016"= "2016-05", "2017" = "2017-05", "2018" = "2018-05")

NationalHomeSales<- pivot_longer(NationalHomeSales, cols= c("1996","1997", "1998", "1999","2000","2001", "2002", "2003", "2004","2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018"), names_to = "YR", values_to = "ZHVI")

ggplot(NationalHomeSales, aes(YR, ZHVI)) +
  facet_wrap(~Metro) +
  geom_point() +
  labs( title = "City Home Value Comparison", x= "Year", y= "Home Value (In$)") + 
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, size=5)) +
  scale_y_continuous(name="Home Value (In$)", labels = scales::comma)
```

